---
title: "xts: eXtensible Time Series"
---

## Using the xts package

The `xts` package provides an extensible time series class that provides 
uniform handling of many R time series classes by extending zoo.

### Installation

The xts package can be installed from CRAN as follows
```{r eval=FALSE}
install.packages("xts")
```

### Creating data objects: `as.xts` and `xts`

Existing time-series data can be converted to an `xts` object with `as,xts`.
<!-- from ?'as.xts' -->
It is possible with a call to `as.xts` to convert objects of class 
`timeSeries`, `ts`, `irts`, `fts`, `its`, `matrix`, `data.frame`, and `zoo`.
```{r tidy=FALSE}
library(xts)
data(sample_matrix)
class(sample_matrix)
matrix_xts <- as.xts(sample_matrix, dateFormat='Date')
str(matrix_xts)
df_xts <- as.xts(as.data.frame(sample_matrix),
                 important='very important info!')
str(df_xts)
```

Data objects can also be constructed directly from raw data with the `xts` 
constructor function, in essentially the same way a `zoo` object is created 
with the exception that at present there is no equivelant `zooreg` class.
```{r}
xts(1:10, Sys.Date()+1:10)
```

### Subsetting
<!-- from ?'[.xts' -->
One of the primary motivations, and key points of differentiation of the 
time series class xts, is the ability to subset rows by specifying ISO-8601 
compatible range strings. This allows for natural range-based time queries 
without requiring prior knowledge of the underlying time object used in 
construction.

When a raw character vector is used for the i subset argument, it is processed 
as if it was ISO-8601 compliant. This means that it is parsed from left to 
right, according to the following specification:

```
CCYYMMDD HH:MM:SS.ss+
```
A full description will be expanded from a left-specified truncated one.

Additionally, one may specify range-based queries by simply supplying two time 
descriptions seperated by a forward slash:

```
CCYYMMDD HH:MM:SS.ss+/CCYYMMDD HH:MM:SS.ss
```

Extract a specific date
```{r}
matrix_xts['2007-01-06']
```

Extract the entire month of March
```{r}
matrix_xts['2007-03']
```

Extract all the data from the beginning through January 7, 2007.
```{r}
matrix_xts['/2007-01-07']
```

Extract all the data from June 25, 2007 through the end.
```{r}
matrix_xts['2007-06-25/']
```

Extract data from May 28, 2007 through June 2, 2007
```{r}
matrix_xts['2007-05-28/2007-06-02']
```

see `help("[.xts")` for more details and examples for extracting subsets of 
`xts` objects.

### About

**Authors:** Jeffrey A. Ryan and [Joshua M. Ulrich](https://github.com/joshuaulrich)

[xts on CRAN](http://cran.r-project.org/web/packages/xts/index.html)

[xts on github](https://github.com/joshuaulrich/xts)

